Flatiron cleaning

### 2021-11-17 Rmd/rdi analysis.Rmd
Update

### step 1: Patients with a date of drug or date of surgery BEFORE diagnosis will be excluded
```{r drugs episode cleanup}
drugs1 <- drugs %>% 
  filter(episodedatasource == "Administrations" & 
           ismaintenancetherapy == "FALSE") %>%
  # 1. 60 days rule between diagnosis and treatment # Loss 1
  # bind with clinical
  inner_join(., clinical_data, #%>% select("patientid", "diagnosisdate", "surgerydate"),
          by = "patientid") %>% 
  arrange(patientid, episodedate) %>% 
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "months")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "months")) %>% 
  ungroup()
```
We have `r drugs1 %>% filter(days_at_first_drug < 0) %>% distinct(patientid) %>% nrow()` patients with date of drug before diagnosis and `r drugs1 %>% filter(days_at_surg < 0) %>% distinct(patientid) %>% nrow()` patients with a date of surgery before diagnosis. Patients left :

@@

  # 3. Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
# There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.
# For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.
# Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend. 
# Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate, drugname, units, surgerydate, days_at_first_drug, days_at_surg, diagnosisdate) %>%
  summarise_at(vars(amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  ungroup() %>% 
  mutate(across(starts_with("amount"), ~ na_if(., "NA"))) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  mutate(across(starts_with("amount"), ~ as.numeric(.))) %>% 
  # mutate(amount_1 = as.numeric(amount_1), amount_2 = as.numeric(amount_2),
  #        amount_3 = as.numeric(amount_3), amount_4 = as.numeric(amount_4)) %>%
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

@@

### stwp " remove patients with adj too far from sutgery
```{r}
remove_delayed_drug <- therapy_line %>% 
  mutate(interval = case_when(
    chemotherapy_type == "Adjuvant" & 
      linenumber_adj == 1       ~interval(start = surgerydate, end = linestartdate_new) / duration(n= 1, units = "months")),
  ) %>% select(patientid, episodedate, surgerydate, diagnosisdate, linenumber_new, linenumber_adj, interval, therapy, chemotherapy_type) %>% arrange(patientid, episodedate) %>% 
  filter(interval > 3) %>% 
  distinct(patientid, .keep_all = TRUE)
remove_delayed_drug <- paste(remove_delayed_drug$patientid, collapse = "|")

therapy_line <- therapy_line %>% 
  # Remove patients with more than 1 line as Neo
  filter(!str_detect(patientid, remove_delayed_drug))
therapy_line %>% distinct(patientid) %>% nrow()

@@

<!-- `r emo::ji("question_mark")`   -->
<!-- <span style="color:red">**Random question : within 1 cycle we can have taxanes multiple times, is it possible for carboplatin or cisplatin?**</span>   -->

@@

  arrange(patientid, episodedate) %>% 
  group_by(patientid, linenumber_new) %>%
  # Need the number of time of the "major" drug is seen in the line
  mutate(cycle_count_perline = min(drugname_count_perline)) %>%
  mutate(cycle_drugname = case_when(
    cycle_count_perline == drugname_count_perline          ~ cycle_drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  group_by(patientid, linenumber_new, cycle_drugname) %>%
  mutate(cycle_increment = row_number(cycle_drugname)) %>% # Look at F002DD2953889 had a long time without one drug------------------
  # F01D908A013B7 has more pacli at the end but in the same line

@@

  # group_by(patientid, linenumber_new, cycle_inc) %>%
  # mutate(cycle_count_1 = row_number(cycle_inc)) %>% 
  # mutate(cycle_c = ifelse(cycle_count_1 == 1, cycle_inc, NA_real_)) %>% 
  arrange(patientid, episodedate, cycle_increment) %>% 
  group_by(patientid, linenumber_new) %>%
  # mutate(cycle_increment = cycle_inc) %>% 
  fill(cycle_increment, .direction = "downup") %>%  # make sure I can do up for have one new drug compare to the linename FBB5186A15CD9, FE593EDF5586A
  group_by(patientid, chemotherapy_type, cycle_increment) %>% 
  mutate(is_dose_dense = row_number(cycle_increment),
         is_dose_dense = ifelse(is_dose_dense == 1, 1, NA_real_)) %>% 
  group_by(patientid) %>% 
  mutate(total_cycle_increment = row_number(is_dose_dense)) %>% 
  fill(total_cycle_increment, .direction = "downup") %>% 
  # cycle date
  group_by(patientid, linenumber_new, cycle_increment) %>% 
  mutate(cycle_start_date = min(episodedate)) %>% 
  # select(patientid, linename, linenumber_new, drugname, episodedate,
  #        cycle_count_perline, cycle_drugname, drugname_count_perline, cycle_increment, cycle_start_date) %>%

  # # CODE FOR SKIPPED CYCLE
  # group_by(patientid, linenumber_new) %>% 
  # mutate(cycle_interval = cycle_start_date - lag(cycle_start_date), 
  #        cycle_interval = ifelse(cycle_interval == 0, NA_real_, cycle_interval)) %>%  # FDB58481AFFBB 48 days
  # 
  # # take for granted that the cycle interval days are the routine cycle for the rest of the line FAACDC7205803
  # mutate(n = row_number(drugname)) %>%
  # 
  # mutate(base_days_between_cycle = ifelse(n==2 , cycle_interval, NA_real_)) %>%
  # fill(base_days_between_cycle, .direction = "downup") %>% 
  # mutate(skipped_cycle_indays = ifelse(((cycle_interval - base_days_between_cycle) != 0) ,
  #                                      (cycle_interval - base_days_between_cycle), NA_real_)) %>% # weird FDB58481AFFBB
  # # Fix skipped_cycle_indays when only 1 drug FDB58481AFFBB skipped_cycle_indays = NA, no other choice...
  # group_by(patientid, linenumber_new) %>%
  # mutate(drug_count_perline = n()) %>%
  # mutate(skipped_cycle_indays1 = ifelse(
  #   (cycle_count_perline == drug_count_perline), NA_real_, skipped_cycle_indays 
  # )) %>% # F001443D8B85C Pb with when different pattern by cycle => May not be able to have the skipped cycle
  # # or need to compare with dose like if 100 should be 7 days and if 200 should be 14 days....
  # group_by(patientid) %>% 
  # mutate(mean_skipped_cycle_indays = mean(skipped_cycle_indays, na.rm = TRUE)) %>% 
  ungroup()
therapy_cycle %>% distinct(patientid) %>% nrow()

@@

```{r remove more than 9 cycle patients}
too_much_cycle <- therapy_cycle %>% 
  filter(
    (therapy == "Upfront Surgery" & total_cycle_increment > 9) |
      (therapy == "Upfront Chemo" & chemotherapy_type == "Neoadjuvant" & total_cycle_increment > 9)
  ) %>%
  distinct(patientid)
too_much_cycle <- paste(too_much_cycle$patientid, collapse = "|")

>>>>

### step 1: Patients with a date of drug or date of surgery BEFORE diagnosis will be excluded
```{r drugs episode cleanup}
drugs1 <- drugs %>% 
  filter(episodedatasource == "Administrations" #& 
           # ismaintenancetherapy == "FALSE"
         ) %>%
  # 1. 60 days rule between diagnosis and treatment # Loss 1
  # bind with clinical
  inner_join(., clinical_data, #%>% select("patientid", "diagnosisdate", "surgerydate"),
          by = "patientid") %>% 
  arrange(patientid, episodedate) %>% 
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "months")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "months")) %>% 
  ungroup()
```
We have `r drugs1 %>% filter(days_at_first_drug < 0) %>% distinct(patientid) %>% nrow()` patients with date of drug before diagnosis and `r drugs1 %>% filter(days_at_surg < 0) %>% distinct(patientid) %>% nrow()` patients with a date of surgery before diagnosis. Patients left :

@@

  # 3. Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
# There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.
# For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.
# Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend. 
# Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate, drugname, units, surgerydate, days_at_first_drug, days_at_surg, diagnosisdate, ismaintenancetherapy) %>%
  summarise_at(vars(amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  ungroup() %>% 
  mutate(across(starts_with("amount"), ~ na_if(., "NA"))) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  mutate(across(starts_with("amount"), ~ as.numeric(.))) %>% 
  # mutate(amount_1 = as.numeric(amount_1), amount_2 = as.numeric(amount_2),
  #        amount_3 = as.numeric(amount_3), amount_4 = as.numeric(amount_4)) %>%
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

@@

<!-- ### step : remove patients with adj too far from surgery -->
```{r}
# remove_delayed_drug <- therapy_line %>% 
#   mutate(interval = case_when(
#     chemotherapy_type == "Adjuvant" & 
#       linenumber_adj == 1       ~interval(start = surgerydate, end = linestartdate_new) / duration(n= 1, units = "months")),
#   ) %>% select(patientid, episodedate, surgerydate, diagnosisdate, linenumber_new, linenumber_adj, interval, therapy, chemotherapy_type) %>% arrange(patientid, episodedate) %>% 
#   filter(interval > 3) %>% 
#   distinct(patientid, .keep_all = TRUE)
# remove_delayed_drug <- paste(remove_delayed_drug$patientid, collapse = "|")
# 
# therapy_line <- therapy_line %>% 
#   # Remove patients with more than 1 line as Neo
#   filter(!str_detect(patientid, remove_delayed_drug))
# therapy_line %>% distinct(patientid) %>% nrow()

@@

`r emo::ji("question_mark")`
<span style="color:red">**Patients have maintenance as first line Adjuvant then recurrent treatment : Should I remove these patients or keep them for the Neoadjuvant RDI??**</span>
```{r}
therapy_line %>% 
  filter(linenumber_adj == 1) %>% 
  distinct(patientid, linenumber_new, .keep_all = TRUE) %>% 
  select(ismaintenancetherapy, linename) %>% 
  tbl_summary(by = linename)
```

<!-- <span style="color:red">**Random question : within 1 cycle we can have taxanes multiple times, is it possible for carboplatin or cisplatin?**</span>   -->

@@

  arrange(patientid, episodedate) %>% 
  group_by(patientid, linenumber_new) %>%
  # Need the number of time of the "major" drug is seen in the line
  mutate(cycle_count_perline = min(drugname_count_perline)) %>%
  mutate(cycle_major_drugname = case_when(
    cycle_count_perline == drugname_count_perline          ~ cycle_drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  group_by(patientid, linenumber_new, cycle_major_drugname) %>%
  mutate(cycle_increment = row_number(cycle_major_drugname)) %>% # Look at F002DD2953889 had a long time without one drug------------------
  # F01D908A013B7 has more pacli at the end but in the same line

@@

  # group_by(patientid, linenumber_new, cycle_inc) %>%
  # mutate(cycle_count_1 = row_number(cycle_inc)) %>% 
  # mutate(cycle_c = ifelse(cycle_count_1 == 1, cycle_inc, NA_real_)) %>% 
  arrange(patientid, episodedate, cycle_increment) %>% 
  group_by(patientid, linenumber_new) %>%
  # mutate(cycle_increment = cycle_inc) %>% 
  fill(cycle_increment, .direction = "downup") %>%  # make sure I can do up for have one new drug compare to the linename FBB5186A15CD9, FE593EDF5586A
  group_by(patientid, chemotherapy_type, cycle_increment) %>% 
  mutate(is_dose_dense = row_number(cycle_increment),
         is_dose_dense = ifelse(is_dose_dense == 1, 1, NA_real_)) %>% 
  group_by(patientid) %>% 
  mutate(total_cycle_increment = row_number(is_dose_dense)) %>% 
  fill(total_cycle_increment, .direction = "downup") %>% 
  # cycle date
  group_by(patientid, linenumber_new, cycle_increment) %>% 
  mutate(cycle_start_date = min(episodedate)) %>%

  # Look at dose dense
  group_by(patientid, linenumber_new, cycle_increment, cycle_drugname) %>%  # weird dose dense F064D9EB733B3 F050B815FC126
  mutate(dose_dense = n()) %>% 
  mutate(dose_dense = max(dose_dense)) %>%  # too much cycle
  ungroup() # %>% 
  # select(patientid, linename, linenumber_new, drugname, episodedate, chemotherapy_type, cycle_increment, dose_dense,
  #        is_dose_dense,cycle_major_drugname, cycle_drugname,
  #        cycle_count_perline, cycle_major_drugname, drugname_count_perline, cycle_start_date) #%>%  # F002DD2953889, F003F2BEEDB37 only 1+2

  # # CODE FOR SKIPPED CYCLE
  # group_by(patientid, linenumber_new) %>% 
  # mutate(cycle_interval = cycle_start_date - lag(cycle_start_date), 
  #        cycle_interval = ifelse(cycle_interval == 0, NA_real_, cycle_interval)) %>%  # FDB58481AFFBB 48 days
  # 
  # # take for granted that the cycle interval days are the routine cycle for the rest of the line FAACDC7205803
  # mutate(n = row_number(drugname)) %>%
  # 
  # mutate(base_days_between_cycle = ifelse(n==2 , cycle_interval, NA_real_)) %>%
  # fill(base_days_between_cycle, .direction = "downup") %>% 
  # mutate(skipped_cycle_indays = ifelse(((cycle_interval - base_days_between_cycle) != 0) ,
  #                                      (cycle_interval - base_days_between_cycle), NA_real_)) %>% # weird FDB58481AFFBB
  # # Fix skipped_cycle_indays when only 1 drug FDB58481AFFBB skipped_cycle_indays = NA, no other choice...
  # group_by(patientid, linenumber_new) %>%
  # mutate(drug_count_perline = n()) %>%
  # mutate(skipped_cycle_indays1 = ifelse(
  #   (cycle_count_perline == drug_count_perline), NA_real_, skipped_cycle_indays 
  # )) %>% # F001443D8B85C Pb with when different pattern by cycle => May not be able to have the skipped cycle
  # # or need to compare with dose like if 100 should be 7 days and if 200 should be 14 days....
  # group_by(patientid) %>% 
  # mutate(mean_skipped_cycle_indays = mean(skipped_cycle_indays, na.rm = TRUE)) %>% 
  # ungroup()
therapy_cycle %>% distinct(patientid) %>% nrow()

@@

# Remove too much cycle
```{r remove more than 9 cycle patients}
therapy_cycle %>% 
  filter(linenumber_new == 1) %>% 
  arrange(patientid, linenumber_new, desc(dose_dense)) %>% 
  mutate(dose_dense = as.factor(dose_dense)) %>%  # too much cycle
  distinct(patientid, linenumber_new, .keep_all = TRUE) %>% 
  select(dose_dense, linename) %>% 
  tbl_summary(by = linename)

too_much_cycle <- therapy_cycle %>%
  filter(
    (therapy == "Upfront Surgery" & total_cycle_increment > 9) |
      (therapy == "Upfront Chemo" & chemotherapy_type == "Neoadjuvant" & total_cycle_increment > 9)
  ) %>%
  distinct(patientid)
too_much_cycle <- paste(too_much_cycle$patientid, collapse = "|")
